Supplementary Material for \\Continuous gravitational waves in the lab: recovering audio signals with a table-top optical microphone

### 2021-10-05 paper/ifo-appendix.tex
grammar and spelling

Two important points must be made about the above procedure. 
(i) Fundamentally the goal is to maximize $\mathcal{L}$ in Eqn.~\ref{eqn:likeOne} by varying $s(t)$ through $A$.
For the special case of the Gaussian likelihood (Eqn.~\ref{eqn:ipnn}), which peaks at $\ip{n}{n}= 0$, this is equivalent to minimizing the distance between $x(t)$ and $s(t)$ as evident in Eqn.~\ref{eqn:ipuv}.
In general, however, minimizing the distance between $x(t)$ and $s(t)$ is not equivalent always to the fundamental goal of maximising $\mathcal{L}$, for example if $\mathcal{L}$ peaks at $\ip{n}{n} \neq 0$, or if $\mathcal{L}$ has multiple maxima.
(ii) The maximum likelihood $\mathcal{L}_{\rm max}$ in Eqn.~\ref{eqn:maxLogLike} (or equivalently its logarithm) defines the detection statistic. 
When its value exceeds a threshold (chosen freely by the analyst) at some value of $\omega$, a signal is deemed to have been deteccted at that value of $\omega$.
Therefore the specific functional form of Eqn.~\ref{eqn:maxLogLike} matters, which is a second reason why one must start from Eqn.~\ref{eqn:ipnn} rather than Eqn.~\ref{eqn:likeOne}, in addition to reason (i).

>>>>

Two important points must be made about the above procedure. 
(i) Fundamentally the goal is to maximize $\mathcal{L}$ in Eqn.~\ref{eqn:likeOne} by varying $s(t)$ through $A$.
For the special case of the Gaussian likelihood (Eqn.~\ref{eqn:ipnn}), which peaks at $\ip{n}{n}= 0$, this is equivalent to minimizing the distance between $x(t)$ and $s(t)$ as evident in Eqn.~\ref{eqn:ipuv}.
In general, however, minimizing the distance between $x(t)$ and $s(t)$ is not equivalent always to the fundamental goal of maximising $\mathcal{L}$, for example if $\mathcal{L}$ peaks at $\ip{n}{n} \neq 0$, or if $\mathcal{L}$ has multiple maxima.
(ii) The maximum likelihood $\mathcal{L}_{\rm max}$ in Eqn.~\ref{eqn:maxLogLike} (or equivalently its logarithm) defines the detection statistic. 
When its value exceeds a threshold (chosen freely by the analyst) at some value of $\omega$, a signal is deemed to have been detected at that value of $\omega$.
Therefore the specific functional form of Eqn.~\ref{eqn:maxLogLike} matters, which is a second reason why one must start from Eqn.~\ref{eqn:ipnn} rather than Eqn.~\ref{eqn:likeOne}, in addition to reason (i).

### 2021-10-05 paper/ifo-conclusions.tex
grammar and spelling

In this paper, we use a table-top Michelson interferometer as an analog to a gravitational-wave detector, demonstrating signal processing techniques used within the gravitational-wave community.
We explore the use of the interferometer as an optical microphone and consider a more general treatment of signal processing with complex audio signals, which can also serve as a distant analog to minimally gravitational-wave burst signals, e.g., from supernovae.
% justify ``un-modeled''? - james
% fair point, it's the lingo used in LIGO, but really it's more accurate to say minimally modelled - Han
The demonstration can be adapted for use in both the physics and engineering undergraduate laboratory, providing opportunities for cross-disciplinary teaching. 
Additionally, it can be used as a tool for explaining gravitational-wave research to a wider, non-specialist audience.

>>>>

In this paper, we use a table-top Michelson interferometer as an analog to a gravitational-wave detector, demonstrating signal processing techniques used within the gravitational-wave community.
We explore the use of the interferometer as an optical microphone and consider a more general treatment of signal processing with complex audio signals, which can also serve as a distant analog to minimally modelled gravitational-wave burst signals, e.g., from supernovae.
% justify ``un-modeled''? - james
% fair point, it's the lingo used in LIGO, but really it's more accurate to say minimally modelled - Han
The demonstration can be adapted for use in both the physics and engineering undergraduate laboratory, providing opportunities for cross-disciplinary teaching. 
Additionally, it can be used as a tool for explaining gravitational-wave research to a wider, non-specialist audience.

### 2021-10-05 paper/ifo-futureWork.tex
grammar and spelling

The experiment described here can be used to demonstrate and teach a variety of topics, from basic optics and photodiode circuits, to signal processing and speech enhancement. 
Cross-disciplinary undergraduate laboratory experiments between physics and engineering courses have been shown to be beneficial to student learning (see Section II in the Supplementary Material for further information). 
Here, we consider further extensions and improvements to the experiment such as noise isolation, transfer function analysis, speech recognition, and control systems.

@@

More traditional techniques such as a wavelet transform~\citep{nason1995stationary} could be used to extract the signal from noise, and compared with the above methods. 
A wavelet transform provides both time and frequency information, making it easier to pinpoint the origin of noise with respect to time. 
In Refs.~\cite{tufekci2000feature,agbinya1996discrete}, wavelet transform methods are proposed for speech recognition.

Returning to the demonstration of gravitational wave analysis, single interferometers cannot yield directional information for signals; a large proportion of the directional information in gravitational-wave detections comes from the time delay offset between the signal\han{note} being recorded at two or more detectors.\cite{GW150914}
We could therefore extend this analysis to include data from two interferometers to extract directional information.
This would require increasing the sensitivity of the optical microphone to pick up the signal from a distant source instead of from a speaker attached directly to one of the mirrors of the interferometer.
Another extension would be to demonstrate the Doppler effect of the Earth's motion around the Sun, which needs to be considered in continuous-wave searches (see Section~\ref{sec:realCWSearches} and Ref.~\cite{JKS:1998}). 
One approach could be to modify the input audio signal to simulate Doppler modulation. 
% did we cut the turn-table idea? that was fun! - james
% yeah afraid so - it was nice, but the editor did not like it! :( - Hannah

>>>>

The experiment described here can be used to demonstrate and teach a variety of topics, from basic optics and photodiode circuits, to signal processing and speech enhancement. 
Cross-disciplinary undergraduate laboratory experiments between physics and engineering courses are beneficial to student learning (see Section II in the Supplementary Material for further information). 
% ``have been shown to be beneficial'' changed to ``are beneficial'' since we justify it elsewhere - james
Here, we consider further extensions and improvements to the experiment such as noise isolation, transfer function analysis, speech recognition, and control systems.

@@

More traditional techniques such as a wavelet transform~\citep{nason1995stationary} could be used to extract the signal from noise and compared with the above methods. 
A wavelet transform provides both time and frequency information, making it easier to pinpoint the origin of noise with respect to time. 
In Refs.~\cite{tufekci2000feature,agbinya1996discrete}, wavelet transform methods are proposed for speech recognition.

Returning to the demonstration of gravitational wave analysis, single interferometers cannot yield directional information for signals; a large proportion of the directional information in gravitational-wave detections comes from the time delay offset between the signal\han{note} being recorded at two or more detectors.\cite{GW150914}
We could therefore extend this analysis to include data from two interferometers to extract directional information.
This would require increasing the sensitivity of the optical microphone to pick up the signal from a distant source instead of from a speaker attached directly to one of the mirrors of the interferometer.
Another extension would be to demonstrate the Doppler effect of the Earth's motion around the Sun, which needs to be considered in continuous-wave searches (see Section~\ref{sec:realCWSearches} and Ref.~\cite{JKS:1998}). 
One approach could be to modify the input audio signal to simulate Doppler modulation. 
% did we cut the turn-table idea? that was fun! - james
% yeah afraid so - it was nice, but the editor did not like it! :( - Hannah
% haha, okay. - james

### 2021-10-05 paper/ifo-supplementary-material.tex
grammar and spelling

The experimental apparatus presented here provides an opportunity to bring together physics and electrical engineering topics in the lab environment. 
Several examples of cross-disciplinary teaching can be found in the literature, for example see Refs.~\cite{OConnerSibrayForinash:2001, PiersonGurlandCrawford:2002, CafarelliEtAl:2012, MakanKopaszGingl:2014}. 
Cross-disciplinary experimental projects between physics and engineering students can increase student understanding of overlapping concepts through contextual learning and also improve student motivation~\cite{PiersonGurlandCrawford:2002}. 
Realistic and open-ended projects spanning both topics allow students a better understanding of real-world application of these topics and of the dependency between physicists and engineers (as shown by the collaborative project presented in Ref.~\cite{OConnerSibrayForinash:2001}).
The optical microphone presented here can similarly be used as a teaching tool for students to conduct investigations into signal processing techniques with a physical system.

>>>>

The experimental apparatus presented here provides an opportunity to bring together physics and electrical engineering topics in the lab environment. 
Several examples of cross-disciplinary teaching can be found in the literature, for example, see Refs.~\cite{OConnerSibrayForinash:2001, PiersonGurlandCrawford:2002, CafarelliEtAl:2012, MakanKopaszGingl:2014}. 
Cross-disciplinary experimental projects between physics and engineering students can increase student understanding of overlapping concepts through contextual learning and also improve student motivation~\cite{PiersonGurlandCrawford:2002}. 
Realistic and open-ended projects spanning both topics allow students a better understanding of real-world applications of these topics and of the dependency between physicists and engineers (as shown by the collaborative project presented in Ref.~\cite{OConnerSibrayForinash:2001}).
The optical microphone presented here can similarly be used as a teaching tool for students to conduct investigations into signal processing techniques with a physical system.
